### 2018-10-17 Paper/examples.tex
..

%
Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ 
enables construction of the FSS with minimal effort. FSS predictions at
a large set of MC samples in the input space are used to obtain converged
estimates of $\mathcal{T}(\theta_i)$. Shown in
Figure~\ref{fig:sense_bore}~(left) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ MC
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/sense_borehole}
  \includegraphics[width=0.51\textwidth]{./Figures/ub_conv_borehole}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain
parameters in the
borehole discharge function in~\eqref{eq:bore}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), plotted
against number of samples. Also included in the legend are estimates of
$\mathcal{T}(\theta_i)$ in each case in the legend.}
\label{fig:sense_bore}
\end{center}
\end{figure}

In Figure~\ref{fig:sense_bore}~(right), we plot estimates of the screening parameter 
$\widehat{\mathcal{C}_i\mu_i}$ for a wide range of the number of 
samples used for approximating $\mu_i$ using~\eqref{eq:mu}.
Estimates for $\widehat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement
with $\mathcal{T}(\theta_i)$ even when small number of samples (5--10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent 
with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear 
that the discharge is insensitive to $T_u$ and $T_l$. 
Moreover, the sensitivity to $K_w$ is also small. We exploit these findings to reduce
the dimensionality of the problem: we can 
discount the variabilities in $T_u$, $T_l$, and $K_w$ by fixing 
them at their respective nominal values.

Our goal as discussed earlier is to gain computational advantage by constructing
surrogates in a reduced input parameter space. To this end, we use LAR to 
construct PCEs
in 5D and 4D spaces by fixing $\{T_u,T_l\}$ in the former and additionally
fixing $K_w$ in the latter at their respective mean values. In
Figure~\ref{fig:conv_bore}~(left), we compare convergence of PCEs constructed in the
full space (7D) with those constructed in the two reduced spaces (4D and 5D)
using $\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_borehole}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_borehole}
\caption{Left: A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. Right: A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}
%
As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}~(left)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}.

As discussed earlier in this section, the reduced-space PCE's are verified for
predictive accuracy in a least-squares sense and a probabilistic sense.
Estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ based on 50 samples in the
validation test suite were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's, respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. Note that $\epsilon_\text{\tiny{LOO}}$
however, is lower in the case of 4D PCE (Figure~\ref{fig:conv_bore}).
This illustrates the
trade-off between accuracy and computational efficiency for the present 
problem. Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced-space surrogates based on the ranking of the parameter sensitivities.

@@

Figure~\ref{fig:conv_bore}~(right) illustrates a comparison of the PDFs

@@

We aim to construct a reduced-space surrogate for the following QoI:
\be
\mathcal{F}(\vec\theta) = \frac{1}{|D|} \int_D u(\vec{x}; \vec\theta) \, d\vec{x}, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite differences, and Newton's method
was used to solve the resulting nonlinear system on a $100 \times 100$ 2D
cartesian grid.
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left) using FSS predictions at 10$^6$
MC samples in the input space. The FSS was 
constructed using model predictions at 500 training points in the 10-dimensional
input space. Corresponding value of $\epsilon_\text{\tiny{LOO}}$ was 
found to be 9.729$\times 10^{-4}$. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an RSS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the RSS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the RSS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\vec\theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-space PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}

Once again, we verify the accuracy of the RSS by estimating $\epsilon_{\mbox{\tiny{L-2}}}$
using model evaluations at 1000 independent MC samples in the 10D parameter
space. The RSS was found to be accurate
within 5$\%$. In order to bolster confidence in the RSS, we compare PDFs of the QoI
as well as a normalized histogram plot based on sparse model evaluations in the  
validation test-suite, in 
Figure~\ref{fig:conv_elliptic}~(right). While the two PDFs are in favorable agreement,
the modal estimate and the spread in the QoI based on the histogram is also captured
by them. Hence, the RSS could be used with a reasonable degree of confidence to
quantify the uncertainty in $\mathcal{F}(\vec\theta)$ thereby leading to
a computational advantage in this case.

>>>>

%
Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ 
enables construction of the FSS with minimal effort. FSS predictions at
a large set of MC samples in the input space are used to obtain converged
estimates of $\mathcal{T}(\theta_i)$. Shown in
Figure~\ref{fig:sense_bore}(a) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ MC
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

\begin{figure}[htbp]
 \begin{center}
 \begin{tabular}{cc}
  \includegraphics[width=0.45\textwidth]{./Figures/sense_borehole}
  &
  \includegraphics[width=0.49\textwidth]{./Figures/ub_conv_borehole}
  \\ (a) & (b)
  \end{tabular}
  \\
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_borehole2}
  \\ (c)
\caption{
(a) Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain
parameters in the
borehole discharge function in~\eqref{eq:bore}; (b)
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), plotted
against number of samples. Also included in the legend are estimates of
$\mathcal{T}(\theta_i)$ in each case in the legend; \rebut{and (c) Estimates of
$\widehat{\mathcal{C}_i\mu_i}$ for individual parameters using $N$ = 5, 10,
15, and 20 samples in the full parameter space.}}
\label{fig:sense_bore}
\end{center}
\end{figure}

In Figure~\ref{fig:sense_bore}(b), we plot estimates of the screening parameter 
$\widehat{\mathcal{C}_i\mu_i}$ for a wide range of the number of 
samples used for approximating $\mu_i$ using~\eqref{eq:mu}. \rebut{Additionally,
to illustrate a comparison of the estimates of $\widehat{\mathcal{C}_i\mu_i}$ obtained
at the end of the initial screening step, i.e. $N$~=~5, with subsequent steps, we plot
them for individual parameters at $N$~=`5, 10, 15, and 20 samples in Figure~\ref{fig:sense_bore}(c).}
Estimates for $\widehat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement
with $\mathcal{T}(\theta_i)$ even when small number of samples (5--10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent 
with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear 
that the discharge is insensitive to $T_u$ and $T_l$. 
Moreover, the sensitivity to $K_w$ is also small. We exploit these findings to reduce
the dimensionality of the problem: we can 
discount the variabilities in $T_u$, $T_l$, and $K_w$ by fixing 
them at their respective nominal values.

Our goal as discussed earlier is to gain computational advantage by constructing
surrogates in a reduced input parameter space. To this end, we use LAR to 
construct PCEs
in 5D and 4D spaces by fixing $\{T_u,T_l\}$ in the former and additionally
fixing $K_w$ in the latter at their respective mean values. In
Figure~\ref{fig:conv_bore}(a), we compare convergence of PCEs constructed in the
full space (7D) with those constructed in the two reduced spaces (4D and 5D)
using $\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}).

\begin{figure}[htbp]
 \begin{center}
 \begin{tabular}{cc}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_borehole}
  &
  \includegraphics[width=0.45\textwidth]{./Figures/errL2_samples_borehole}
  \\ (a) & (b)
  \end{tabular}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_borehole}
  \\ (c)
\caption{(a) A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. \rebut{The degree of
the resulting PCE was found to be 3, 2,and 5 in the 4D, 5D, and 7D case respectively;
(b) Semilog plot of the relative L-2 error norm ($\epsilon_{\mbox{\tiny{L-2}}}$) 
for the PCEs constructed in 4D, 5D, and 7D};
and (c) A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}
%
As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}(a)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}. 
\rebut{However, the plot of $\epsilon_{\mbox{\tiny{L-2}}}$ as a function of sample
size for the three surrogate models in Figure~\ref{fig:conv_bore}(b)
clearly demonstrates the trade-off between
computational effort associated with surrogate construction and its predictive 
accuracy. For a small sample size of 20, the relative  error in the case of 4D
PCE is found to be the smallest. However, as the sample size increases, the higher
order PCE becomes more accurate. Note that the same set of 50 samples in the
validation test suite were used in each case. Hence, when limited computational resources
are available, the proposed methodology could yield substantial gains due to fast
convergence as well as higher accuracy of the surrogate in the reduced space.
The gains are expected to increase exponentially as the dimensionality of the 
input space increases.}

Specifically, the estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's, respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. Note that $\epsilon_\text{\tiny{LOO}}$
however, is lower in the case of 4D PCE (Figure~\ref{fig:conv_bore}).
Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced-space surrogates based on the ranking of the parameter sensitivities.

@@

Figure~\ref{fig:conv_bore}(c) illustrates a comparison of the PDFs

@@

We aim to construct a reduced-space surrogate for the following QoI:
\be
\mathcal{F}(\vec\theta) = \frac{1}{|D|} \int_D u(\vec{x}; \vec\theta) \, d\vec{x}, 
\label{eq:qoi}
\ee
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite differences, and Newton's method
was used to solve the resulting nonlinear system on a $100 \times 100$ 2D
cartesian grid.
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left) using FSS predictions at 10$^6$
MC samples in the input space. The FSS was 
constructed using model predictions at 500 training points in the 10-dimensional
input space. Corresponding value of $\epsilon_\text{\tiny{LOO}}$ was 
found to be 9.729$\times 10^{-4}$. 
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an RSS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic}(a). 
As expected, the RSS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the RSS in the present test problem
is expected to be much smaller. 
\begin{figure}[htbp]
 \begin{center}
 \begin{tabular}{cc}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_elliptic}
  &
  \includegraphics[width=0.45\textwidth]{./Figures/errL2_samples_elliptic}
  \\ (a) & (b)
  \end{tabular}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_elliptic}
  \\ (c)
\caption{(a) Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample 
size for PCEs constructed in 10 and 5 dimensions to compare their convergence 
characteristics. \rebut{The degree of
the resulting PCE was found to be 3 in both cases; 
(b) Semilog plot of the relative L-2 error norm ($\epsilon_{\mbox{\tiny{L-2}}}$) 
for the PCEs constructed in 5D and 7D}; 
and (c) PDF of the QoI, $\mathcal{F}(\vec\theta)$ in~\eqref{eq:semilinear}
is plotted using the full-surrogate and the reduced-space PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}
\rebut{The comparison of convergence characteristics of $\epsilon_{\mbox{\tiny{L-2}}}$
illustrated in Figure~\ref{fig:conv_elliptic}(b) indicates that for a small sample size of 20,
the surrogate constructed in the reduced space performs better in terms of
predictability. However, as the sample size increases, the full space is observed to
be slightly more accurate. However, the order of the accuracy is found to be 
consistent in both cases until the sample size is less than 80. When the sample size
increases to 90, the 19D PCE is observed to be more accurate by an order of magnitude.
However, the corresponding value of $\epsilon_{\mbox{\tiny{LOO}}}$ is~$\mathcal{O}(10^{-1})$.
Hence, the 19D PCE does not seem to have fully  converged and requires additional 
model evaluations. Therefore, considering
fast convergence of the surrogate in the 5D space and predictive accuracy comparable
to the 19D PCE,
the proposed methodology significantly reduces computational effort in this case.}

Once again, we verify the accuracy of the RSS by estimating $\epsilon_{\mbox{\tiny{L-2}}}$
using model evaluations at 1000 independent MC samples in the 10D parameter
space. The RSS was found to be accurate
within 5$\%$. In order to bolster confidence in the RSS, we compare PDFs of the QoI
as well as a normalized histogram plot based on sparse model evaluations in the  
validation test-suite, in 
Figure~\ref{fig:conv_elliptic}(c). While the two PDFs are in favorable agreement,
the modal estimate and the spread in the QoI based on the histogram is also captured
by them. Hence, the RSS could be used with a reasonable degree of confidence to
quantify the uncertainty in $\mathcal{F}(\vec\theta)$ thereby leading to
a computational advantage in this case.
